Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure

### 2017-04-06 documents/TriangleScenarioArXiv/ts.tex
First pass through at completing the numerical optimizations result section

    \subsection{Noisy Distributions}
    \label{sec:noise}
    \begin{figure}
    \begin{center}
            \includegraphics[scale=0.6,trim={0 0 0 0.4in},clip]{../../figures/distributions/uniform_dist_plot.pdf}
            \vspace{-0.2in}
            \[ \probplotvalue{255, 255, 109} = \f{1}{64} \]
            \caption{The completely uniform distribution $\s{U}$ visualized using a $4 \times 4 \times 4$ grid.}
            \label{fig:uniform_distribution}
    \end{center}
    \end{figure}
    In his original work, \citet{Fritz_2012} was able to find quantum non-classicality in the Triangle structure by embedding Bell non-classicality. As was mentioned in \cref{sec:fritz_distribution}, this embedding required a very idealistic condition to hold: perfect correlations between $C$ and $A_l, B_l$. From an experimental point of view, it is impossible to use Fritz's original argument to confirm non-classicality because every observable distribution is subject to \text{noise}. It is possible to minimize noise by developing more accurate measurement channels but perfect correlations are too idealistic. Causal compatibility inequalities not only permit there to be noise within a set of observations, but they also provide a means of quantifying the amount of noise that is allowed before non-classicality breaks down.

    For Triangle structure, we define a completely noisy observation as the uniform probability distribution $\s{U}\br{abc} = \s{U}_{ABC}\br{abc} = \f{1}{64} \quad \forall a,b,c\in \bc{0,1,2,3}$ (see \cref{fig:uniform_distribution}). Since the uniform distribution $\s{U}$ is trivially compatible with the Triangle structure, the objective of this section is to determine how much uniform noise needs to be added to the Fritz distribution $\prob[\fritz]$ before it transitions from incompatible to compatible. We do this by defining an affine noise parameter $\vep$ such that $0 \leq \vep \leq 1$ along with the \term{noisy variant} $\s{N}_{\vep}$ of the Fritz distribution,
    \[ \s{N}_{\vep} = \br{1 - \vep} \prob[\fritz] + \vep \s{U} \]
    Therefore as $\vep$ varies from $0$ to $1$, more and more noise is added to the Fritz distribution and $\s{N}_{\vep}$ transitions from an incompatible distribution $\s{N}_{0} = \prob[\fritz]$ to a compatible distribution $\s{N}_{1} = \s{U}$. For each inequality $I$ and distribution $\prob$, let $I\br{\prob}$ denote a scalar value representing the amount of violation seen by $\prob$.  Our tests suggest that inequality $I_1$ (\cref{eq:web_inequality}) is most robust to noise and demonstrates that the Fritz distribution remains incompatible with the Triangle structure up to at least a noise parameter of $\vep \simeq 0.085$. The associated distribution $\s N_{0.085}$ is plotted in \cref{fig:noisy_fritz}. Of course, there remains the possibility that another inequality will be able to withstand a larger degree of noise than $I_1$.

@@

    By parameterizing unitary matrices, it becomes possible to parameterize $d$-dimensional density matrices and $d$-element PVMs by recognizing that any orthonormal basis $\bc{\ket{\psi_j}}$ (where $1 \leq j \leq d$) can be transformed into the computational basis $\bc{\ket{j}}$ by a unitary transformation $U$, i.e. $U \ket{\psi_j} = \ket{j}$. First consider a $d$-element PVM $M = \bc{\ket{m_j}\bra{m_j} \mid 1 \leq j \leq d}$. Since $\bc{\ket{m_j}}$ forms an orthonormal basis, one can parameterize $M$ by writing $M = \bc{U^{\dagger}\ket{j}\bra{j}U \mid 1 \leq j \leq d}$ and parameterizing $U$ using \cref{eq:spengler_unitary}. This method was inspired by the measurement seeding method for iterative optimization used by \citet{Pal_2010}. Analogously, this argument can be extended to full-rank $d$-dimensional density matrices $\rho$ by performing a spectral decomposition $\rho = \sum_{j=1}^{d} p_j \ket{p_j} \bra{p_j}$ into eigenvalues $\bc{p_j}$ and eigenstates $\bc{\ket{p_j}}$. Since $\Tr\bc{\rho} = \sum_{j = 1}^{d} p_j = 1$ the eigenvalues of $\rho$ are parameterized without degeneracy using a tuple of $d-1$ real-valued parameters with periodicity $\bs{0, 2 \pi}$ using hyper-spherical coordinates~\cite{Hedemann_2013, Spengler_2010_Unitary}. Additionally, since $\rho$ is Hermitian, the eigenstates $\bc{\ket{p_j}}$ form an orthonormal basis and therefore the eigenstates are analogously parameterized using \cref{eq:spengler_unitary}: $\rho = \sum_{j=1}^{d} p_j U^{\dagger} \ket{j} \bra{j} U$. For our purposes, we have set $d = 4$ and fixed $\la_{l, l} = 0$ for $1 \leq l \leq d$ in \cref{eq:spengler_unitary} because the global phase contributions are irrelevant for \cref{eq:quantum_model_triangle_pvms}.

@@

    To quantify the degree in which each inequality $I$ could be further violated, we define a relative violation $V_F$ to be the maximum violation relative to the violation achieved by the Fritz distribution, i.e. $V_F \defined \min_{P}\bc{I\br{P}}/I\br{P_F}$. The first inequality in consideration is $I_1$; the non-symmetric Web inflation inequality presented in \cref{eq:web_inequality}. At best, numerical optimizations converged to a distribution (\cref{fig:maximum_violation_I_1}) with relative violation $V_F \simeq 1.501$. Almost immediately, it is evident that \cref{fig:maximum_violation_I_1} closely resembles the Fritz distribution. In fact, \cref{fig:maximum_violation_I_1} and \cref{fig:fritz_distribution_visualized} share the exact same \textit{possibilistic structure} or set of possible events. What does this reveal about $I_1$ and non-classicality in the Triangle structure? This result is not entirely unexpected. As was mentioned in \cref{sec:found_inequalities}, these inequalities are derived specifically to show incompatibility of the Triangle structure. Consequently, it seems that $I_1$ only has the capacity to demonstrate incompatibility for a specific forms of non-classicality that resemble the Fritz distribution very closely.

@@

    Since the inequalities themselves are polynomial and the space of quantum-accessible distributions is non-convex, a number of optimization methods were used in an attempt to eliminate any issues associated with local minima or ill-conditioned convergence. Specifically, the BFGS method~\cite[p. 142]{Nocedal_2000} and the Nelder-Mead simplex method~\cite[p. 238]{Nocedal_2000} were used along with a method called Basin Hopping~\cite{Wales_1997}, which is a hybrid between simulated annealing and gradient descent-based methods.

    An alternative explanation has to do with the sensitively of the initial condition $\la_{\br{0}} = \br{\la_{\br{0}, 0}, \ldots, \la_{\br{0}, n}}$ in our optimization. In \cref{sec:param_quantum_dist}, it was shown that our parameter space of choice was $2\pi$-periodic in all entries. Somewhat surprisingly, when using a uniform seed for $\la_0$, all optimization methods consistently converged to inequality saturation instead of inequality violation which of course is achievable by the Fritz distribution. Instead, only initial seeds close enough to the \textit{Fritz parameters} $\la_{\br{\fritz}}$ (the parameters that induce the Fritz distribution using the quantum implementation in \cref{sec:fritz_distribution}) where able to achieve violation. To illustrate this graphically, we let the initial seed be a small Gaussian perturbation $\de \la$ away from $\la_{\br{\fritz}}$ in parameter space.
    \[ \la_{\br{0}} = \la_{\br{\fritz}} + \de \la  \]
    The Gaussian perturbation was given zero mean $\mu = 0$ and variance $\si^2 = \br{2 \pi 10^{\chi}}^2$ where $\chi$ measures an order of magnitude standard deviation. For $I_1$, deviations of $\chi \gtrsim 1.3$ inhibited convergence to violation. This is depicted in \cref{fig:initial_conditions}. Therefore, it is unclear whether or not our results share the same possibilistic structure as $\prob[\fritz]$ because of the inequality itself or the ill-conditioned nature of the optimization. Most likely, this phenomena is due to a combination of both effects.

    \begin{figure}
    \begin{subfigure}[t]{.48\linewidth}
    {\centering\includegraphics[width=\linewidth]{../../figures/optimizations/Gaussian_Perturbation_Fritz_Color_Default.pdf}}
    \caption{For each iteration step $k$, the inequality target decreases.}\label{fig:initial_conditions_a}
    \end{subfigure}
    \begin{subfigure}[t]{.48\linewidth}
    {\centering\includegraphics[width=\linewidth]{../../figures/optimizations/Gaussian_Perturbation_Fritz_Color_Zoomed.pdf}}
    \caption{\Cref{fig:noise} zoomed to show convergence is sensitive to initial conditions.}\label{fig:initial_conditions_b}
    \end{subfigure}
    \caption{Numerical optimizations aimed at violating causal compatibility inequalities are sensitive to initial conditions.}\label{fig:initial_conditions}
    \end{figure}

    The second inequality in consideration is $I_3$; the \textit{symmetric} Web inflation inequality presented in \cref{eq:symmetic_web_inequality}. This inequality was able to achieve a greater relative violation than $I_1$ as optimizations converged to a distribution (\cref{fig:maximum_violation_I_3}) with relative violation $V_F \simeq 2.61$. Just as before, the maximum violating distribution shares a possibilistic structure with the Fritz distribution. However, this result is far more surprising. Since the inequality is symmetric with respect to the exchange of parties $\varphi \in \Perm{A, B, C}$, one would naively expect that the maximum violating distribution is be symmetric as well. If it were not, one could construct a distribution that violates $I_3$ more by symmetrizing $\prob[\fritz]$ by summing over all possible permutations and normalizing. The symmetrized Fritz distribution is plotted in \cref{fig:symmetrized_fritz}. This distribution is also non-classical with relative violation $V_F = 3$. As it turns out, the symmetrized Fritz distribution is \textit{not} quantum-accessible in the sense that is cannot be written as \cref{eq:quantum_model_triangle}\footnote{At least numerical optimizations suggest this is the case.} Therefore we conclude that the space of quantum accessible correlations on the Triangle structure is non-convex; a reiteration of a result found in~\citet{Inflation}.
    \begin{figure}
    \begin{center}
            \includegraphics[scale=0.6,trim={0 0 0 0.4in},clip]{../../figures/distributions/plotted_dist_I_3_max_violation_2017.pdf}
            \caption{Probability distribution that maximizes violation of $I_3$ from \cref{eq:symmetic_web_inequality} ($V_F \simeq 2.61$).}
            \label{fig:maximum_violation_I_3}
    \end{center}
    \end{figure}
    \begin{figure}
    \begin{center}
            \includegraphics[scale=0.6,trim={0 0 0 0.4in},clip]{../../figures/distributions/symmetrized_fritz.pdf}
            \caption{Symmetrized version of the Fritz distribution ($V_F \simeq 3$).}
            \label{fig:symmetrized_fritz}
    \end{center}
    \end{figure}

    Finally for both $I_0$ and $I_2$, the maximum relative violation happens to be $V_F = 1$ (up to numerical precision). This means that the Fritz distribution either represents a local minimum of the parameter space or is in fact \textit{the} maximally violating distribution. Without further investigations, it is unclear if this behavior is related to the universality of the Fritz distribution or if its an artifact of the methods used to derive $I_0$ and $I_2$.

>>>>

    \subsection{Noisy Distributions}
    \label{sec:noise}
    % \begin{figure}
    % \begin{center}
    %         \includegraphics[scale=0.6,trim={0 0 0 0.4in},clip]{../../figures/distributions/uniform_dist_plot.pdf}
    %         \vspace{-0.2in}
    %         \[ \probplotvalue{255, 255, 109} = \f{1}{64} \]
    %         \caption{The completely uniform distribution $\s{U}$ visualized using a $4 \times 4 \times 4$ grid.}
    %         \label{fig:uniform_distribution}
    % \end{center}
    % \end{figure}
    In his original work, \citet{Fritz_2012} was able to find quantum non-classicality in the Triangle structure by embedding Bell non-classicality. As was mentioned in \cref{sec:fritz_distribution}, this embedding required a very idealistic condition to hold: perfect correlations between $C$ and $A_l, B_l$. From an experimental point of view, it is impossible to use Fritz's original argument to confirm non-classicality because every observable distribution is subject to \text{noise}. It is possible to minimize noise by developing more accurate measurement channels but perfect correlations are too idealistic. Causal compatibility inequalities not only permit there to be noise within a set of observations, but they also provide a means of quantifying the amount of noise that is allowed before non-classicality breaks down.

    For Triangle structure, we define a completely noisy observation as the uniform probability distribution $\s{U}\br{abc} = \s{U}_{ABC}\br{abc} = \f{1}{64} \quad \forall a,b,c\in \bc{0,1,2,3}$. Since the uniform distribution $\s{U}$ is trivially compatible with the Triangle structure, the objective of this section is to determine how much uniform noise needs to be added to the Fritz distribution $\prob[\fritz]$ before it transitions from incompatible to compatible. We do this by defining an affine noise parameter $\vep$ such that $0 \leq \vep \leq 1$ along with the \term{noisy variant} $\s{N}_{\vep}$ of the Fritz distribution,
    \[ \s{N}_{\vep} = \br{1 - \vep} \prob[\fritz] + \vep \s{U} \]
    Therefore as $\vep$ varies from $0$ to $1$, more and more noise is added to the Fritz distribution and $\s{N}_{\vep}$ transitions from an incompatible distribution $\s{N}_{0} = \prob[\fritz]$ to a compatible distribution $\s{N}_{1} = \s{U}$. For each inequality $I$ and distribution $\prob$, let $I\br{\prob}$ denote a scalar value representing the amount of violation seen by $\prob$.  Our tests suggest that inequality $I_1$ (\cref{eq:web_inequality}) is most robust to noise and demonstrates that the Fritz distribution remains incompatible with the Triangle structure up to at least a noise parameter of $\vep \simeq 0.085$. The associated distribution $\s N_{0.085}$ is plotted in \cref{fig:noisy_fritz}. Of course, there remains the possibility that another inequality will be able to withstand a larger degree of noise than $I_1$.

@@

    By parameterizing unitary matrices, it becomes possible to parameterize $d$-dimensional density matrices and $d$-element PVMs by recognizing that any orthonormal basis $\bc{\ket{\psi_j}}$ (where $1 \leq j \leq d$) can be transformed into the computational basis $\bc{\ket{j}}$ by a unitary transformation $U$, i.e. $U \ket{\psi_j} = \ket{j}$. First consider a $d$-element PVM $M = \bc{\ket{m_j}\bra{m_j} \mid 1 \leq j \leq d}$. Since $\bc{\ket{m_j}}$ forms an orthonormal basis, one can parameterize $M$ by writing $M = \bc{U^{\dagger}\ket{j}\bra{j}U \mid 1 \leq j \leq d}$ and parameterizing $U$ using \cref{eq:spengler_unitary}. This method was inspired by the measurement seeding method for iterative optimization used by \citet{Pal_2010}. Analogously, this argument can be extended to full-rank $d$-dimensional density matrices $\rho$ by performing a spectral decomposition $\rho = \sum_{j=1}^{d} p_j \ket{p_j} \bra{p_j}$ into eigenvalues $\bc{p_j}$ and eigenstates $\bc{\ket{p_j}}$. Since $\Tr\br{\rho} = \sum_{j = 1}^{d} p_j = 1$ the eigenvalues of $\rho$ are parameterized without degeneracy using a tuple of $d-1$ real-valued parameters with periodicity $\bs{0, 2 \pi}$ using hyper-spherical coordinates~\cite{Hedemann_2013, Spengler_2010_Unitary}. Additionally, since $\rho$ is Hermitian, the eigenstates $\bc{\ket{p_j}}$ form an orthonormal basis and therefore the eigenstates are analogously parameterized using \cref{eq:spengler_unitary}: $\rho = \sum_{j=1}^{d} p_j U^{\dagger} \ket{j} \bra{j} U$. For our purposes, we have set $d = 4$ and fixed $\la_{l, l} = 0$ for $1 \leq l \leq d$ in \cref{eq:spengler_unitary} because the global phase contributions are irrelevant for \cref{eq:quantum_model_triangle_pvms}.

@@

@@

    To quantify the degree in which each inequality $I$ could be further violated, we define a relative violation $V_F$ to be the maximum violation relative to the violation achieved by the Fritz distribution, i.e. $V_F \defined \min_{P}\bc{I\br{P}}/I\br{P_F}$. The first inequality in consideration is $I_1$; the non-symmetric Web inflation inequality presented in \cref{eq:web_inequality}. At best, numerical optimizations converged to a distribution (\cref{fig:maximum_violation_I_1}) with relative violation $V_F \simeq 1.501$. Almost immediately, it is evident that \cref{fig:maximum_violation_I_1} closely resembles the Fritz distribution. In fact, \cref{fig:maximum_violation_I_1} and \cref{fig:fritz_distribution_visualized} share the exact same \textit{possibilistic structure}, i.e. the same set of possible events. This result is not entirely unexpected. As was mentioned in \cref{sec:found_inequalities}, these inequalities are derived specifically to prove the incompatibility of the Fritz distribution and the Triangle structure. Consequently, it seems that $I_1$ only has the capacity to demonstrate incompatibility for distributions that resemble the Fritz distribution very closely.

    An alternative explanation for qualitative similarities between \cref{fig:maximum_violation_I_1} and \cref{fig:fritz_distribution_visualized} is related to the sensitively of the initial parameters to our optimization. Above it was shown that our parameter space of choice was $2\pi$-periodic in all entries. Somewhat surprisingly, when supplied with randomly sampled initial parameters, all optimization methods consistently converged to inequality \textit{saturation} instead of inequality \textit{violation} which is achievable by the Fritz distribution $\prob[\fritz]$. Instead, it was observed that only parameters close enough to the \textit{Fritz parameters} (the parameters that induce the Fritz distribution using the quantum implementation in \cref{sec:fritz_distribution}) were able to any violation. Since the inequalities themselves are polynomial and the space of quantum-accessible distributions is non-convex, a number of optimization methods were used in an attempt to eliminate any issues associated with local minima or ill-conditioned convergence. Specifically, the BFGS method~\cite[p. 142]{Nocedal_2000} and the Nelder-Mead simplex method~\cite[p. 238]{Nocedal_2000} were used along with a method called Basin Hopping~\cite{Wales_1997}, which is a hybrid between simulated annealing and gradient descent-based methods. Despite our varied efforts, it remains unclear whether or not our results share the same possibilistic structure as $\prob[\fritz]$ because of the inequality itself or the ill-conditioned nature of the optimization. More likely, this phenomena is due to a combination of both effects.

    % The Gaussian perturbation was given zero mean $\mu = 0$ and variance $\si^2 = \br{2 \pi 10^{\chi}}^2$ where $\chi$ measures an order of magnitude standard deviation. For $I_1$, deviations of $\chi \gtrsim 1.3$ inhibited convergence to violation. This is depicted in \cref{fig:initial_conditions}.

    % \begin{figure}
    % \begin{subfigure}[t]{.48\linewidth}
    % {\centering\includegraphics[width=\linewidth]{../../figures/optimizations/Gaussian_Perturbation_Fritz_Color_Default.pdf}}
    % \caption{For each iteration step $k$, the inequality target decreases.}\label{fig:initial_conditions_a}
    % \end{subfigure}
    % \begin{subfigure}[t]{.48\linewidth}
    % {\centering\includegraphics[width=\linewidth]{../../figures/optimizations/Gaussian_Perturbation_Fritz_Color_Zoomed.pdf}}
    % \caption{\Cref{fig:noise} zoomed to show convergence is sensitive to initial conditions.}\label{fig:initial_conditions_b}
    % \end{subfigure}
    % \caption{Numerical optimizations aimed at violating causal compatibility inequalities are sensitive to initial conditions.}\label{fig:initial_conditions}
    % \end{figure}

    The second inequality in consideration is the \textit{symmetric} inequality $I_3$ ($\cref{eq:symmetic_web_inequality}$). This inequality was able to achieve a greater relative violation than $I_1$ as optimizations converged to a distribution (\cref{fig:maximum_violation_I_3}) with relative violation $V_F \simeq 2.61$. Just as before, the maximum violating distribution shares a possibilistic structure with the Fritz distribution. However, this result is far more surprising. Since the inequality is symmetric with respect to the permutation of parties, one would naively expect that the maximum violating distribution be symmetric as well. If it were not, one could construct a distribution that violates $I_3$ more by symmetrizing $\prob[\fritz]$ by summing over all possible permutations and normalizing. The symmetrized Fritz distribution is plotted in \cref{fig:symmetrized_fritz}. This distribution is also non-classical with relative violation $V_F = 3$. As it turns out, numerical optimizations suggest that the symmetrized Fritz distribution is \textit{not} quantum-accessible in the sense that is cannot be written as \cref{eq:quantum_model_triangle}. Therefore we conclude that the space of quantum accessible correlations on the Triangle structure is non-convex; a reiteration of a result mentioned in~\citet{Inflation}.
    \begin{figure}
    \begin{center}
            \includegraphics[scale=0.6,trim={0 0 0 0.4in},clip]{../../figures/distributions/plotted_dist_I_3_max_violation_2017.pdf}
            \caption{Probability distribution that maximizes violation of $I_3$ from \cref{eq:symmetic_web_inequality} ($V_F \simeq 2.61$).}
            \label{fig:maximum_violation_I_3}
    \end{center}
    \end{figure}
    \begin{figure}
    \begin{center}
            \includegraphics[scale=0.6,trim={0 0 0 0.4in},clip]{../../figures/distributions/symmetrized_fritz.pdf}
            \caption{Symmetrized version of the Fritz distribution ($V_F \simeq 3$).}
            \label{fig:symmetrized_fritz}
    \end{center}
    \end{figure}

    Finally for both $I_0$ and $I_2$, the maximum relative violation happens to be $V_F = 1$ (up to numerical precision). This means that the Fritz distribution either represents a local minimum of the parameter space or is in fact \textit{the} maximally violating distribution. Without further investigations, it remains unclear if this behavior is related to the universality of the Fritz distribution or if its an artifact of the methods used to derive $I_0$ and $I_2$.

    \subsection{Rosset Variation}
    \label{sec:rosset_variation}
    \todo[TC]{This might take a page}
